---
title: "Diamonds data sets"
author: "Fatimah Alamer"
date: "10/1/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,eval = TRUE)
```
## Introduction 

Diamonds data sets is one of popular example data sets in R 
lets gets familiar with data set 
**first step** we have to install tidyvers package
```{r , echo = TRUE, eval = FALSE}
install.packages("tidyverse")
```
**second step**: Initialize package 

```{r , echo = TRUE, eval = TRUE}
library(tidyverse)
```
Now, our diamonds data sets 
```{r , echo = TRUE, eval = TRUE}
diamonds
```


Now, lets gets familiar with data, you may think to use summary, or we can use glimpse function from tidyvers package, you can now see the difference

```{r  eval = TRUE}
summary(diamonds)


glimpse(diamonds)
```

Now, we know some information regarding the data:
+There are 53,940 round-cut diamonds,that is exactly the number of rows!
+There are 10 variables:cut, color, & clarity,carat, depth, table, x, y, z, price

## Applying some function in tidyvers package
### 1-data transformation with dplyr
 now,we want to know the price regarding for each clarity type:


```{r }

diamonds%>%
  group_by(clarity ) %>%
  summarise(avg = mean(price),
            mimum= min(price),
            max= max(price),
            stdev = sd(price))
```
for more usefull information,  see [dplyr cheat sheet](https://nyu-cdsc.github.io/learningr/assets/data-transformation.pdf)

   
### 2-Data tidying with tidyr
  lets make our data tidy.
```{r }
diamonds%>%
  
  pivot_longer(c(x,y,z), names_to = "dimension", values_to = "value") -> diamonds_tidy

           
```
after making tidy data long , we have 161820 observation ! increasing number of observation will makes it easier to reproduce analyses because they are more easy to understand, update, and reuse.
for more useful information,  see [tidyer cheat sheet](https://github.com/rstudio/cheatsheets/blob/master/tidyr.pdf')

### 3-data vizulization with ggplot2**
lets have the first ggplot:
 
```{r }

ggplot(diamonds_tidy, aes(carat, price)) + geom_point()+
geom_point(shape = 16, alpha = 0.5, position = "jitter")
```


```{r}
ggplot(diamonds_tidy, aes(carat, price,color = clarity)) + 
geom_point(shape = 16, alpha = 0.5, position = "jitter")+
      geom_smooth(method=lm, se=FALSE)+
facet_grid(cut~.)+
 labs(x = "carat",
       y = "price",
       title = "The diamonds Dataset",
       color = "clarity")

```





For more useful information,  see [ggplot2 cheat sheet](https://nyu-cdsc.github.io/learningr/assets/data-visualization-2.1.pdf)

## 4-Table:the largest 20 diamonds in the `diamonds` dataset
```{r results = "asis"}
library(knitr)
diamonds %>%
  arrange(desc(carat)) %>%
  slice(1:20) %>%
  select(carat, cut, color, clarity) %>%
  knitr::kable(
    caption = "The largest 20 diamonds in the `diamonds` dataset."
  )
```
